The human in the loop: volunteer computing as a socio-technical system

### 2016-02-28 Alife16/paper.tex
Intro

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time they will
``lend''. This makes them an essential part of the algorithm and of
the performance of the whole system. As a sociotechnical system, this
participation follows some patterns and in this paper we examine the
result of several volunteer distributed evolutionary computation
experiments and try to find out which are those patterns and what
makes an experiment successful or not, including the feedback loop
that is created between the users and the algorithm itself.
\end{abstract}

@@

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate islands that
carry out their own evolutionary algorithms. This is why NodIO is
proposed mainly as a {\em complement} to existent resources such as
desktop systems or laptops. As long as it provides a non-null
computational capability that can help existing resources find the
solution faster it will have found its purpose. The main use case is
someone setting up NodIO in the cloud, writing a fitness evaluation
function and running a client from his or her own computer, but
requesting help in social networks for additional resources. That is
why, in this system,  we have considered the whole {\em social} aspect
in the design, with issues related to security, trust and privacy
among others. The computing system becomes a {\em techno-social
  system} \citep{vespignani2009predicting}. In this paper we are going
to focus on measuring the response of users to experiments, but at the
same time we will also focus on the technical aspects of the server and how
those might change the behavior of users, improving the capability of
the system.

>>>>

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other resources they will
``lend''. This makes them an essential part of the algorithm and of
the performance of the whole system. As a sociotechnical system, this
participation follows some patterns and in this paper we examine the
result of several volunteer distributed evolutionary computation
experiments and try to find out which are those patterns and what
makes an experiment successful or not, including the feedback loop
that is created between the users and the algorithm itself.
\end{abstract}

@@

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island 
carrying out their own evolutionary algorithms. This is why NodIO is
proposed mainly as a {\em complement} to existent resources such as
desktop systems or laptops. As long as it provides a non-null
computational capability that can help existing resources find the
solution faster it will have found its purpose. The main use case is
someone setting up NodIO in the cloud, writing a fitness evaluation
function and running a client from his or her own computer, but
requesting help in social networks for additional resources. That is
why, in this system,  we have considered the whole {\em social} aspect
in the design, with issues related to security, trust and privacy
among others. The computing system becomes a {\em techno-social
  system} \citep{vespignani2009predicting}. In this paper we are going
to focus on measuring the response of users to experiments, but at the
same time we will also focus on the technical aspects of the server and how
those might change the behavior of users, improving the capability of % Mario: How these might change
the system.
